---
title: "Colapse"
author: "Mason Carhart"
date: "4/9/2021"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(pacman)
p_load(data.table, dplyr, collapse, microbenchmark, xaringan, kniter, nycflights13, lubridate)
```

# Collapse Package

## Perks

- Increased efficiency 

- Simple implementation 

- Uses C/C++


---

# Speed Comparison and Motivation 
```{r Speed Test 1, echo=TRUE}

storms_dt = as.data.table(storms)

microbenchmark(collapse = storms_dt %>% gby(category) %>% get_vars(10:11) %>% fmean,
               data.table = storms_dt[, lapply(.SD, mean), keyby = category
                                             , .SDcols = 10:11], times = 100)

```

---
# Unfair Test

```{r Speed Test 2, echo=TRUE}

storms_dt = as.data.table(storms)

microbenchmark(collapse = storms_dt %>% gby(category) %>% get_vars(10:11) %>% fmean,
                      data.table = storms_dt[, lapply(.SD, mean),
                                             .SDcols = 10:11], times = 100)

```

---

# Collapse


- dplyr integration 

```{r, echo=TRUE}

storms_dt %>% group_by(year) %>% select_at(6:7) 
            %>% fmean %>% tail(4)# Fast Mean of wind and pressure

```

---
- data.table integration

All the code has been done from a data.table object.

```{r, echo=TRUE}

storms_dt %>% fsubset( year, wind, pressure) %>% # Fast Fillter
  ftransform(wind_over_presure_per = wind / pressure * 100,  # Computing % of wind to pressure
             wind_mean = fmean(wind)) %>%    # Mean wind
             tail()

```

---
- Time series and other class integration

```{r, echo=TRUE}

storms_fselected <- storms_dt %>% fselect(year, wind, pressure) %>% na_omit

storms_lubridate <- storms_dt %>% ftransform(date = make_date(year, month, day)) %>%  
  fselect(date, wind, pressure) %>% na_omit %>% head()



```


---
# Bigger Data Set

```{r Speed Test 3, echo=TRUE}

sflights_dt = as.data.table(flights) 

microbenchmark(collapse = sflights_dt %>% fgroup_by(year) %>% 
                    get_vars(15:16) %>% fmean,
               data.table = sflights_dt[, lapply(.SD, mean), keyby = year
                    , .SDcols = 15:16], times = 100)

```

---

# Helpfull Rescorces and Links

- Collapse Git Hub Page https://github.com/SebKrantz/collapse

- Function Overview https://sebkrantz.github.io/collapse/reference/index.html

- Collapse Cheat Sheet https://raw.githubusercontent.com/SebKrantz/cheatsheets/master/pngs/collapse.png






